A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-10 sections/Metatheory.tex
measurements -> measures

We have proven this theorem by a lexicographic group of induction measurements\\
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
on the worklist $\Gm$.
$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of teams
and the total ``difficulty'' of judgments, respectively;
$|\cdot|_\forall$ and $|\cdot|_\to$ count the total number of
universal quantifiers and function types, respectively;
$|\cdot|_{\al}$ counts the number of existential variables in the worklist.

@@

It is not difficult to see that all but two algorithmic reduction rules
decreases the group of measurements.
(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measurements.)
The two exceptions are Rules 10 and 11; they increases the number of existential variables
without decreasing the number of universal quantifiers.
However, at the same time, they both break the subtyping problem into two sub-problems
(Rule 7 applied right after Rule 10 or 11), in smaller sizes.
Therefore we argue that such split makes the worklist easier, by analyzing a subset of worklist,
which we call \emph{instantiation judgments} $\Gm_i$, with the help of $\Gm_\le$ for the proof.

@@

Shown in Figure~\ref{fig:worklist_ext},
worklist extension formally describe the minor changes to worklist during algorithmic reduction.
When the algorithm processes the worklist, the first judgment is processed,
and the rest of the worklist roughly remains its shape.
Worklist extension is formally defined by two cases:
existential variable solving ($\mathtt{\jExt solve}$) and insertion ($\mathtt{\jExt \al}$).
Existential variable instantiation, as another common form of worklist manipulation,
i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
can be derived with the combination of the 3 rules.
Within expectation, some worklist measurements are not changed modulo worklist extension.

@@

In summary, the decidability theorem is shown through a lexicographic group of induction measurements
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$.
The critical case is that, whenever we encounter an instantiation judgment at the top of the worklist,
we make use of Lemma~\ref{lemma:inst:decidable}, which reduces $|\Gm|_{\al}$ or $|\Gm|_\forall$
by consuming that instantiation judgment. Other cases are relatively straightforward.

>>>>

We have proven this theorem by a lexicographic group of induction measures\\
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
on the worklist $\Gm$.
$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of teams
and the total ``difficulty'' of judgments, respectively;
$|\cdot|_\forall$ and $|\cdot|_\to$ count the total number of
universal quantifiers and function types, respectively;
$|\cdot|_{\al}$ counts the number of existential variables in the worklist.

@@

It is not difficult to see that all but two algorithmic reduction rules
decreases the group of measures.
(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measures.)
The two exceptions are Rules 10 and 11; they increases the number of existential variables
without decreasing the number of universal quantifiers.
However, at the same time, they both break the subtyping problem into two sub-problems
(Rule 7 applied right after Rule 10 or 11), in smaller sizes.
Therefore we argue that such split makes the worklist easier, by analyzing a subset of worklist,
which we call \emph{instantiation judgments} $\Gm_i$, with the help of $\Gm_\le$ for the proof.

@@

Shown in Figure~\ref{fig:worklist_ext},
worklist extension formally describe the minor changes to worklist during algorithmic reduction.
When the algorithm processes the worklist, the first judgment is processed,
and the rest of the worklist roughly remains its shape.
Worklist extension is formally defined by two cases:
existential variable solving ($\mathtt{\jExt solve}$) and insertion ($\mathtt{\jExt \al}$).
Existential variable instantiation, as another common form of worklist manipulation,
i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
can be derived with the combination of the 3 rules.
Within expectation, some worklist measures are not changed modulo worklist extension.

@@

In summary, the decidability theorem is shown through a lexicographic group of induction measures
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$.
The critical case is that, whenever we encounter an instantiation judgment at the top of the worklist,
we make use of Lemma~\ref{lemma:inst:decidable}, which reduces $|\Gm|_{\al}$ or $|\Gm|_\forall$
by consuming that instantiation judgment. Other cases are relatively straightforward.
